---
title: "analysis"
author: "Matthew King-Hang Ma"
date: "2025-08-21"
output: html_document
---

# Import libraries and define useful functions

## Import libraries
```{r}
# For general purpose and plotting
library(Cairo)
library(tidyverse)
library(arrow)
library(ggpubr)
library(ggsci)
library(patchwork)
library(roxygen2)
library(progressr)
handlers(handler_progress(
  format = ":message [:bar] :percent | elapsed: :elapsed | eta: :eta"
))

# For statistical analysis
library(rstatix)
library(emmeans)
library(ez)
library(xtable)
library(lme4)
library(lmerTest)
library(ggstatsplot)
library(MuMIn)
library(glue)
library(here)
```


## Defining relevant functions for ICC computation
These functions are designed to compute the ICC(2,1) from the computed network measurements. 
```{r}

run_foolproof_lmer <- function(formula_str, this_data) {
  # 1. Initialize tracking variables
  captured_logs <- character()
  is_converged = TRUE
  convergence_warning <- FALSE
  is_singular <- FALSE
  
  # 2. Run the model inside handlers to catch "transient" R warnings/messages
  this_lme <- withCallingHandlers(
    lmer(as.formula(formula_str), data = this_data),
    warning = function(w) {
      captured_logs <<- c(captured_logs, w$message)
      invokeRestart("muffleWarning") # Hides it from the console
    },
    message = function(m) {
      # Some lme4 steps use messages instead of warnings
      captured_logs <<- c(captured_logs, m$message)
      invokeRestart("muffleMessage") 
    }
  )
  
  # 3. TARGET A: Check for Singularity
  if (isSingular(this_lme)) {
    is_singular <- TRUE
    captured_logs <- c(captured_logs, "boundary (singular) fit")
  }
  
  # 4. TARGET B: Check Internal lme4 Messages (Convergence)
  internal_msgs <- this_lme@optinfo$conv$lme4$messages
  
  if (!is.null(internal_msgs)) {
    # Add any internal messages (like "Model failed to converge") to our log
    captured_logs <- c(captured_logs, unlist(internal_msgs))
    
    # Flag specifically if it's a convergence issue
    if (any(grepl("failed to converge", internal_msgs, ignore.case = TRUE))) {
      is_converged = FALSE
      convergence_warning <- TRUE
    }
  }
  
  # 5. Clean up the logs (remove empty strings, line breaks, and duplicates)
  captured_logs <- captured_logs[captured_logs != ""]
  captured_logs <- gsub("\n", " ", captured_logs) # Keep logs on a single line
  captured_logs <- unique(trimws(captured_logs))
  
  # 6. Return a clean list with everything you need
  return(list(
    model = this_lme,
    is_singular = is_singular,
    is_converged = is_converged,
    # failed_to_converge = convergence_warning,
    all_warnings = captured_logs
  ))
}


compute_icc_2way_boot = function(df_all, this_network_var, this_band, this_thres_value, this_effcycles, k, 
                      nboot = 1000) {
  this_data = df_all %>% 
    filter(Band == this_band, ThresValue == this_thres_value, EffCycles == this_effcycles)
  
  # Try two-way model first
  formula_str_2way = paste0(this_network_var, " ~ 1 + (1|SubjectCode) + (1|rand_seed)")
  
  # Fit two-way model with catching warnings
  convergence_warning_2way = FALSE
  this_lme_2way = withCallingHandlers(
    lmer(as.formula(formula_str_2way), this_data),
    warning = function(w) {
      convergence_warning_2way <<- TRUE
      warning_message <<- w$message  # Capture the actual warning text
    }
  )
  
  # Check if two-way model is viable
  is_singular_2way = isSingular(this_lme_2way)
  is_converged_2way = !convergence_warning_2way
  
  # If two-way model fails, fit one-way model
  use_oneway = is_singular_2way || !is_converged_2way
  
  if(use_oneway) {
    formula_str = paste0(this_network_var, " ~ 1 + (1|SubjectCode)")
    convergence_warning = FALSE
    this_lme = withCallingHandlers(
      lmer(as.formula(formula_str), this_data),
      warning = function(w) {
        convergence_warning <<- TRUE
      }
    )
    is_singular = isSingular(this_lme)
    is_converged = !convergence_warning
  } else {
    this_lme = this_lme_2way
    is_singular = is_singular_2way
    is_converged = is_converged_2way
  }
  
  # Extract variance components
  vc = lme4::VarCorr(this_lme)
  
  # Extract variances differently based on model type
  if(!use_oneway) {
    # For two-way model
    v_subject = as.numeric(vc$SubjectCode)
    v_seed = as.numeric(vc$rand_seed)
    v_residual = attr(vc, "sc")^2
    
    # Compute ICCs for two-way model
    icc_single = v_subject / (v_subject + v_seed + v_residual)
    icc_k = v_subject / (v_subject + v_seed/k + v_residual/k)
  } else {
    # For one-way model
    v_subject = as.numeric(vc$SubjectCode)
    v_seed = 0  # No seed variance in one-way model
    v_residual = attr(vc, "sc")^2
    
    # Compute ICCs for one-way model
    icc_single = v_subject / (v_subject + v_residual)
    icc_k = v_subject / (v_subject + v_residual/k)
  }
  
  # Bootstrap function
  get_icc <- function(model) {
    vc <- VarCorr(model)
    if(length(vc) > 1) {  # two-way model
      v_subj <- as.numeric(vc$SubjectCode)
      v_seed <- as.numeric(vc$rand_seed)
      v_res <- attr(vc, "sc")^2
      icc1 <- v_subj / (v_subj + v_seed + v_res)
      icck <- v_subj / (v_subj + v_seed/k + v_res/k)
    } else {  # one-way model
      v_subj <- as.numeric(vc$SubjectCode)
      v_res <- attr(vc, "sc")^2
      icc1 <- v_subj / (v_subj + v_res)
      icck <- v_subj / (v_subj + v_res/k)
    }
    c(ICC_2_1 = icc1, ICC_2_k = icck)
  }
  
  # Perform bootstrap if model converged
  if(is_converged) {
    boot_results <- bootMer(this_lme, 
                           FUN = get_icc, 
                           nsim = nboot, 
                           .progress = "txt")
    
    ci_icc1 = quantile(boot_results$t[,1], c(0.025, 0.975))
    ci_icck = quantile(boot_results$t[,2], c(0.025, 0.975))
  } else {
    ci_icc1 = c(NA, NA)
    ci_icck = c(NA, NA)
  }
  
  # Return results
  df_icc = as_tibble(list(
    ICC_2_1 = icc_single,
    ICC_2_k = icc_k,
    ICC_2_1_CI_lower = ci_icc1[1],
    ICC_2_1_CI_upper = ci_icc1[2],
    ICC_2_k_CI_lower = ci_icck[1],
    ICC_2_k_CI_upper = ci_icck[2],
    var_subject = v_subject,
    var_seed = v_seed,
    var_residual = v_residual,
    is_singular = is_singular,
    is_converged = is_converged,
    used_oneway = use_oneway,
    warning_triggered = warning_message %||% NA_character_  # Capture warning message if exists
  ))
  
  return(df_icc)
}

#' Compute ICC with Two-Way Model and Fallback, with Bootstrapped CIs
#'
#' This function calculates the Intraclass Correlation Coefficient (ICC) using a two-way
#' linear mixed-effects model. If the two-way model fails to converge or is singular,
#' it gracefully falls back to a simpler one-way model. It also performs a
#' parametric bootstrap to compute confidence intervals for the ICC values.
#'
#' @param df_all A data frame containing all the data.
#' @param this_network_var A character string of the network variable to be analyzed.
#' @param this_band A character string of the frequency band.
#' @param this_thres_value A numeric threshold value.
#' @param this_effcycles An integer for the number of effective cycles.
#' @param k An integer representing the number of repeated measures per subject.
#' @param nboot An integer for the number of bootstrap replicates (default is 1000).
#' @param ncores An integer for the number of CPU cores to use for parallel bootstrapping (default is 1).
#'
#' @return A tibble with one row containing the calculated ICCs, their bootstrapped
#'         confidence intervals, variance components, and logical flags indicating
#'         model status.
#'
#' @importFrom lme4 lmer VarCorr isSingular bootMer
#' @importFrom dplyr filter
#' @importFrom tibble as_tibble
#' @importFrom stats quantile
compute_icc_2way_boot_parallel = function(df_all, this_network_var, this_band, this_thres_value, this_effcycles, k, nboot = 500, ncores = 1) {
  
  # Filter the data for the specific combination of parameters
  this_data = df_all %>%
    filter(Band == this_band, ThresValue == this_thres_value, EffCycles == this_effcycles)
  #-----------------------------------------------------------------------------
  # Model Fitting
  #-----------------------------------------------------------------------------
  
  # Try to fit the two-way model first
  formula_str_2way = paste0(this_network_var, " ~ 1 + (1|SubjectCode) + (1|rand_seed)")
  
  list_results_2way = run_foolproof_lmer(formula_str_2way, this_data)
  this_lme = list_results_2way$model
  is_singular_2way = list_results_2way$is_singular
  is_converged_2way = list_results_2way$is_converged
  
  # Determine if we need to use the one-way model
  use_oneway = is_singular_2way || !is_converged_2way
  
  if (use_oneway) {
    message(glue("Two-way model failed for {this_network_var}, Band={this_band}, Thres={this_thres_value}, Effcycle={this_effcycles}. Falling back to one-way model."))
    formula_str = paste0(this_network_var, " ~ 1 + (1|SubjectCode)")
    
    list_results_oneway = run_foolproof_lmer(formula_str, this_data)
    this_lme = list_results_oneway$model
    is_singular_oneway = list_results_oneway$is_singular
    message(is_singular_oneway)
    is_converged_oneway = list_results_oneway$is_converged
  } 
  else {
    is_singular_oneway = NA
    is_converged_oneway = NA
  }
  
  #-----------------------------------------------------------------------------
  # Variance Components and ICC Calculation
  #-----------------------------------------------------------------------------
  
  if (use_oneway){
    this_lme = list_results_oneway$model
    is_converged = is_converged_oneway
  } else {
    this_lme = list_results_2way$model
    is_converged = is_converged_2way
  }
  
  vc = lme4::VarCorr(this_lme)
  v_subject = as.numeric(vc$SubjectCode)
  v_residual = attr(vc, "sc")^2
  
  # Check if rand_seed variance component exists
  if ("rand_seed" %in% names(vc)) {
    v_seed = as.numeric(vc$rand_seed)
  } else {
    v_seed = 0
  }
  
  # Compute ICCs
  icc_single = v_subject / (v_subject + v_seed + v_residual)
  icc_k = v_subject / (v_subject + v_seed/k + v_residual/k)
  
  #-----------------------------------------------------------------------------
  # Bootstrapping
  #-----------------------------------------------------------------------------
  
  # Define the function to be bootstrapped
  get_icc <- function(model) {
    vc <- lme4::VarCorr(model)
    v_subj <- as.numeric(vc$SubjectCode)
    v_res <- attr(vc, "sc")^2
    
    # Check if the rand_seed variance exists
    if ("rand_seed" %in% names(vc)) {
      v_seed <- as.numeric(vc$rand_seed)
    } else {
      v_seed <- 0
    }
    
    icc1 <- v_subj / (v_subj + v_seed + v_res)
    icck <- v_subj / (v_subj + v_seed/k + v_res/k)
    
    c(ICC_2_1 = icc1, ICC_2_k = icck)
  }
  
  ci_icc1 = c(NA, NA)
  ci_icck = c(NA, NA)
  
  if (is_converged) {
    boot_results <- tryCatch({
      bootMer(this_lme,
              FUN = get_icc,
              nsim = nboot,
              seed = 20260513,
              parallel = "multicore",
              ncpus = ncores)
    }, error = function(e) {
      warning("Bootstrapping failed for this model.", call. = FALSE)
      NULL
    })
    
    if (!is.null(boot_results)) {
      ci_icc1 = quantile(boot_results$t[, 1], c(0.025, 0.975))
      ci_icck = quantile(boot_results$t[, 2], c(0.025, 0.975))
    }
  }
  
  #-----------------------------------------------------------------------------
  # Return Results
  #-----------------------------------------------------------------------------
  
  df_icc = as_tibble(list(
    ICC_2_1 = icc_single,
    ICC_2_k = icc_k,
    ICC_2_1_CI_lower = ci_icc1[1],
    ICC_2_1_CI_upper = ci_icc1[2],
    ICC_2_k_CI_lower = ci_icck[1],
    ICC_2_k_CI_upper = ci_icck[2],
    var_subject = v_subject,
    var_seed = v_seed,
    var_residual = v_residual,
    is_singular_twoway = is_singular_2way,
    is_singular_oneway = is_singular_oneway %||% NA,  # NA if one-way model was not used
    is_converged_twoway = is_converged_2way,
    is_converged_oneway = is_converged_oneway %||% NA,  # NA if one-way model was not used
    used_oneway = use_oneway
  ))
  return(df_icc)
}

# # A wrapper function of `compute_icc_2way_boot_parallel` to run through all combinations
compute_icc_results_boot = function(df_input, list_network_var, list_band, list_thres_value, list_cycles, k, ncores){
  with_progress({
    p <- progressor(steps = length(list_network_var) * length(list_band) * length(list_thres_value) * length(list_cycles))
    icc_results = crossing(network_var = list_network_var,
                           Band = list_band,
                           ThresValue = list_thres_value,
                           EffCycles = list_cycles,
                           k = k,
                           ncores = ncores) %>%
      group_by(network_var, Band, ThresValue, EffCycles, k) %>%
      group_modify(~ {
        p() # this will increment the progress bar
        compute_icc_2way_boot_parallel(df_input, .y$network_var, .y$Band, .y$ThresValue, .y$EffCycles, .y$k)
      })
  })

  icc_results$Band = factor(icc_results$Band, list_band)
  icc_results$network_var = factor(icc_results$network_var, list_network_var)
  return(icc_results)
}



```


# Data preprocessing

## Data integration: Loading dataframe computed from S5.

To load the dataframes computed from S5.
```{r}
list_band = c("delta", "theta", "alpha", "beta", "gamma")
list_task = c("EyesOpenNoTask", "EyesClosedNoTask")

load_df = function(this_task, this_band){
  dir_derivative = here("derivative", "study_reliability", "group-connectivity", "ciplv")
  fname = list.files(dir_derivative, glue("task-{this_task}.*filt-{this_band}.*"))
  df = arrow::read_feather(file.path(dir_derivative, fname))
  df$EffCycles = as.numeric(df$EffCycles)
  return(df)
}

# They are loaded separately to avoid an extremely large dataframe
df_eo = crossing(BandInit=list_band) %>% group_by(BandInit) %>% group_modify(~load_df("EyesOpenNoTask", .y$BandInit)) %>% ungroup() %>% select(-BandInit)

df_ec = crossing(BandInit=list_band) %>% group_by(BandInit) %>% group_modify(~load_df("EyesClosedNoTask", .y$BandInit)) %>% ungroup() %>% select(-BandInit)

```

## Data cleaning, reduction and integration
```{r}

# Filter out those rows with SWP < 0, which might due to errors in the SWP computation.
df_ec = df_ec %>% filter(SWP > 0)
df_eo = df_eo %>% filter(SWP > 0)

df_all = bind_rows(df_ec, df_eo)

```


# Computing reliability map

Use bootstrapping approach with two-way random intercepts model

## Reliability map: CC
Reliability map for clustering coefficient
```{r}
list_thres_value = df_all$ThresValue %>% unique()

# Computing reliability map for theta+ bands
list_network_var = c("clustering_coef")
list_band = c("theta", "alpha", "beta", "gamma")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 36, 42, 48, 54, 60, 66, 72, 78, 84, 90, 96, 102, 108, 114, 120)
df_ec_cc_icc_results_boot_fourbands = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
# saveRDS(df_ec_cc_icc_results_boot_fourbands, "df_ec_cc_icc_results_boot_fourbands.rds")
saveRDS(df_ec_cc_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_ec_cc_icc_results_boot_fourbands.rds"))

df_eo_cc_icc_results_boot_fourbands = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
# saveRDS(df_eo_cc_icc_results_boot_fourbands, "df_eo_cc_icc_results_boot_fourbands.rds")
saveRDS(df_eo_cc_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_eo_cc_icc_results_boot_fourbands_v20260513.rds"))

# Computing reliability map for delta bands
list_network_var = c("clustering_coef")
list_band = c("delta")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30)
df_ec_cc_icc_results_boot_delta = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_ec_cc_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_ec_cc_icc_results_boot_delta.rds"))

df_eo_cc_icc_results_boot_delta = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_eo_cc_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_eo_cc_icc_results_boot_delta.rds"))
```

## Reliability map: GE
Reliability map for global efficiency
```{r}
list_thres_value = df_all$ThresValue %>% unique()

# Computing reliability map for theta+ bands
list_network_var = c("global_eff")
list_band = c("theta", "alpha", "beta", "gamma")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 36, 42, 48, 54, 60, 66, 72, 78, 84, 90, 96, 102, 108, 114, 120)
df_ec_ge_icc_results_boot_fourbands = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_ec_ge_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_ec_ge_icc_results_boot_fourbands.rds"))

df_eo_ge_icc_results_boot_fourbands = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_eo_ge_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_eo_ge_icc_results_boot_fourbands.rds"))

# Computing reliability map for delta bands
list_network_var = c("global_eff")
list_band = c("delta")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30)
df_ec_ge_icc_results_boot_delta = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_ec_ge_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_ec_ge_icc_results_boot_delta.rds"))

df_eo_ge_icc_results_boot_delta = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_eo_ge_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_eo_ge_icc_results_boot_delta.rds"))
```


## Reliability map: SWP
Reliability map for small-world propensity
```{r}
list_thres_value = df_all$ThresValue %>% unique()

# Computing reliability map for theta+ bands
list_network_var = c("SWP")
list_band = c("theta", "alpha", "beta", "gamma")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 36, 42, 48, 54, 60, 66, 72, 78, 84, 90, 96, 102, 108, 114, 120)
df_ec_swp_icc_results_boot_fourbands = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_ec_swp_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_ec_swp_icc_results_boot_fourbands.rds"))

df_eo_swp_icc_results_boot_fourbands = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_eo_swp_icc_results_boot_fourbands, here("computed_v_EMBC2026Refactor", "df_eo_swp_icc_results_boot_fourbands.rds"))

# Computing reliability for delta band
list_network_var = c("SWP")
list_band = c("delta")
list_cycles = c(6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30)
df_ec_swp_icc_results_boot_delta = compute_icc_results_boot(df_ec, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_ec_swp_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_ec_swp_icc_results_boot_delta.rds"))

df_eo_swp_icc_results_boot_delta = compute_icc_results_boot(df_eo, list_network_var, list_band, list_thres_value, list_cycles, 30, 14)
saveRDS(df_eo_swp_icc_results_boot_delta, here("computed_v_EMBC2026Refactor", "df_eo_swp_icc_results_boot_delta.rds"))
```



